---
title: "Data prep"
author: "Mykola Dereva"
date: "5/18/2021"
output: html_document
---

## Project description

The project was inspired by this [blogpost](https://habr.com/ru/company/yandex_praktikum/blog/557256/) (in russian).

The idea is that they make questionnaire to check how CV photo influence on the subjective perception of candidate competency.

They claim that small smile on candidate's CV photo, slightly increase the perception of candidate competency

However, the blogpost lack of any kind of statistical data analysis and relies only on descriptive statistics.

Luckily they [shared the collected data](https://docs.google.com/spreadsheets/d/1QJbmdpYDQz5mpiqyg08Z5RuP--2q2m893to401wUbi0/edit#gid=41231417) of questionnaire, so I can play with the data on my own and check their findings.

## Load data

```{r}
rm(list = ls())
library(tidyverse)
```

Set locale to display Russian characters properly

```{r}
Sys.setlocale("LC_CTYPE", "russian")
```

```{r}
d <- read_csv("How CV photo influence feel of competence (in Russia).csv")
```

```{r}
head(d)
```

First of all, I need to translate data into english to make the further analysis easier Meanwhile, I will delete the columns I do not need for the analysis

```{r}
d %>% 
  rename(resp_sex = `Укажите свой пол`,        # resp for responder
         resp_age = `Укажите свой возраст`,
         resp_recruting = `Насколько часто вы участвуете в найме людей?`,
         model_id = `Номер участника`,
         model_sex = `Пол участника`,
         model_emotion = `Эмоция`,
         competency_feel = `Оценка`)
```
